### 2016-02-19 DishSimulations/memo/aarondefns.tex
Stuck in merge conflict  Please enter the commit message for your changes. Lines starting

\def\BCCP{$^2$}
\def\BCCPtxt{$^2$Berkeley Center for Cosmological Physics, University of California, Berkeley, Berkeley, California 94720, USA }

\def\ASU{$^7$}
\def\ASUtxt{$^7$Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA}

\def\myemail{\altaffilmark{*}}
\def\myemailtxt{\altaffiltext{*}{aaronew@mit.edu}}

\def\UW{$^6$}
\def\UWtxt{$^6$University of Washington, Department of Physics, Seattle, WA 98195, USA}

\def\SKASA{$^8$}
\def\SKASAtxt{$^8$Square Kilometre Array South Africa (SKA SA), Park Road, Pinelands 7405, South Africa}

\def\RU{$^9$}
\def\RUtxt{$^9$Department of Physics and Electronics, Rhodes University, Grahamstown 6140, South Africa}

\def\CfA{$^3$}
\def\CfAtxt{$^3$Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}

\def\ANU{$^{10}$}
\def\ANUtxt{$^{10}$Australian National University, Research School of Astronomy and Astrophysics, Canberra, ACT 2611, Australia}

\def\CAASTRO{$^{11}$}
\def\CAASTROtxt{$^{11}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}

\def\Haystack{$^{12}$}
\def\Haystacktxt{$^{12}$MIT Haystack Observatory, Westford, MA 01886, USA}

\def\MIT{$^1$}
\def\MITtxt{$^1$MIT Kavli Institute for Astrophysics and Space Research, Cambridge, MA 02139, USA}

\def\Curtin{$^{13}$}
\def\Curtintxt{$^{13}$International Centre for Radio Astronomy Research, Curtin University, Perth, WA 6845, Australia}

\def\Victoria{$^{16}$}
\def\Victoriatxt{$^{16}$Victoria University of Wellington, School of Chemical \& Physical Sciences, Wellington 6140, New Zealand}

\def\UWisc{$^{17}$}
\def\UWisctxt{$^{17}$University of Wisconsin--Milwaukee, Department of Physics, Milwaukee, WI 53201, USA}

\def\UMichigan{$^{18}$}
\def\UMichigantxt{$^{18}$University of Michigan, Department of Atmospheric, Oceanic and Space Sciences, Ann Arbor, MI 48109, USA}

\def\UMelbourne{$^{19}$}
\def\UMelbournetxt{$^{19}$The University of Melbourne, School of Physics, Parkville, VIC 3010, Australia}

\def\USydney{$^{14}$}
\def\USydneytxt{$^{14}$The University of Sydney, Sydney Institute for Astronomy, School of Physics, NSW 2006, Australia}

\def\CASS{$^{20}$}
\def\CASStxt{$^{20}$CSIRO Astronomy and Space Science (CASS), PO Box 76, Epping, NSW 1710, Australia}

\def\Tata{$^{21}$}
\def\Tatatxt{$^{21}$National Centre for Radio Astrophysics, Tata Institute for Fundamental Research, Pune 411007, India}

\def\RRI{$^{22}$}
\def\RRItxt{$^{22}$Raman Research Institute, Bangalore 560080, India}

\def\NRAO{$^{23}$}
\def\NRAOtxt{$^{23}$National Radio Astronomy Observatory, Charlottesville and Greenbank, USA}

\def\UWA{$^{24}$}
\def\UWAtxt{$^{24}$International Centre for Radio Astronomy Research, University of Western Australia, Crawley, WA 6009, Australia}

\def\ASTRON{$^5$}
\def\ASTRONtxt{$^5$Netherlands Institute for Radio Astronomy (ASTRON), PO Box 2, 7990 AA Dwingeloo, The Netherlands}

\def\Dunlap{$^{15}$}
\def\Dunlaptxt{$^{15}$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON M5S 3H4, Canada}

\def\SNS{$^4$}
\def\SNStxt{$^4$Scuola Normale Superiore, Piazza dei Cavalieri 7, I-56126 Pisa, Italy}

\newcommand{\dtb}{T_b}
\newcommand{\Mpci}{\text{ Mpc}^{-1}}
\newcommand{\Mpc}{\text{ Mpc}}
\newcommand{\MHz}{\text{ MHz}}
\newcommand{\GHz}{\text{ GHz}}
\newcommand{\mJy}{\text{ mJy}}
\newcommand{\muJy}{ \text{ } \mu \text{Jy}}
\newcommand{\nJy}{\text{ nJy}}
\newcommand{\Msun}{ M_{\odot}}
\newcommand{\ssection}{$\S$}
\newcommand{\erg}{\text{ erg}}
\newcommand{\second}{\text{ sec}}
\newcommand{\yr}{\text{ yr}}
\newcommand{\degree}{^{\circ}}
\newcommand{\kperp}{k_\perp}
\newcommand{\kpara}{k_\parallel}
\newcommand{\mto}{$\to$}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\mK}{\text{ mK}}
\newcommand{\K}{\text{ K}}
\newcommand{\meter}{\text{ m}}
\newcommand{\km}{\text{km}}
\newcommand{\hfine}{21\,cm}

\def\Let@{\def\\{\notag\math@cr}}